---
title: "a3_task3_aaron_kreisberg"
author: "Aaron Kreisberg"
date: "11/06/2020"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(janitor)
library(here)
library(kableExtra)
library(lubridate)
```

First the urchins.csv file will be read in using read.csv(here("data", "urchins.csv")) from the data folder in the R.project, and names will be cleaned using clean_names(). 

```{r}
urchins <- read.csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

## Section 1 - Exploratory data visualization of purple and red urchins

First the distribution of sizes of purple and red urchins will be examined visually. This will be done by using gg_plot, with both a faceted histogram (geom_histogram + facet_wrap(~common_name)) and a boxplot (geom_boxplot). 

```{r}
ggplot(data = urchins, aes(x = size, fill = common_name)) +
  geom_histogram() +
  facet_wrap(~common_name) +
  labs(x = "Urchin Size (nearest 0.5 cm)", y = "# of Urchins")
```
```{r}
ggplot(data = urchins, aes(x = common_name, y = size, fill = common_name)) +
  geom_boxplot() +
  labs(x = "Urchins", y ="Urchin Size (nearest 0.5 cm)")
 
  
```

Next the data will be explored in a summary statistics table

```{r}
urchin_stats <- urchins %>%
  group_by(common_name) %>% 
  summarise(mean_size = mean(size),
            median_size = median(size),
            sd_size = sd(size), 
            n = n()) 
kbl(urchin_stats) %>% 
  kable_classic(full_width = TRUE)


```
 
Below where your exploratory graphs and table appear, add 2 - 3 sentences briefly describing your overall takeaways about how the size (distribution, central tendency, spread, etc.) compare between the two species of urchins (again, here we are ignoring site and treatment as a variable). Help to contextualize any differences you observe - for example, what is the actual and % difference in mean urchin diameter between the species? 

## Section 2 - Exploration of purple urchin counts by site and year

In a piped sequence of code a subset of the urchins data will be created

```{r}
purple_urchins <- urchins %>%
    mutate(date_new = mdy(date)) %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(year = year(date_new)) %>% 
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef", 
    site == "AQUE" ~ "Arroyo Quemado Reef", 
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"))

purple_urchin_site_year <- purple_urchins %>%
  group_by(year, site_full) %>% 
  count(site_full)
  
purple_urchin_site_year  
```

```{r}
ggplot(data = purple_urchin_site_year, aes(x = year, y = n, color = site_full)) +
  scale_x_continuous(breaks = seq(2008, 2020, 2)) +
  geom_line() +
  geom_point()
   
```

